Sparse spectral methods for partial differential equations on spherical caps

### 2020-11-09 papers/spherical-caps/spherical-caps.tex
Added belapsed image and comment
Typo fixes

\bstodoinline{You mentioned before that you didnt like the use of $i$ being an index thoughout (as in $n,k,i$}. Is this still the case?}

@@

Define $X := (J_x^{(0)})^\top$, $Y := (J_y^{(0)})^\top$, $Z := (J_z^{(0)})^\top$. We can obtain the matrix operator for the variable-coefficient function $v(x,y,z)$ by using the Clenshaw algorithm with matrix inputs as the Jacobi matrices $X, Y, Z$, yielding an operator matrix of the same dimension as the input Jacobi matrices a la the procedure introduced in \cite{olver2019triangle}. We can denote the resulting operator acting on coefficients in the $\bigscopNo$ space by $v(X, Y, Z)$. In matrix-vector notation, our system hence becomes:
\begin{align*}
    (\Delta_W^{(1)} + k^2 \:T^{(0)\to(1)} \: V \: T_W^{(1)\to(0)}) \: \mathbf{u} = \mathbf{f}
\end{align*}
which can be solved to find $\mathbf{u}$. We can see the sparsity and structure of this matrix system in Figure \ref{fig:sparsity} with $v(x,y,z) = zxy^2$ as an example. In Figure \ref{fig:helmholtz} we see the solution to the inhomogeneous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the spherical cap $\Omega$, with $f(x,y,z) = y e^x \genjacw^{(1,0)}(z)$, $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$ and $k = 100$. In Figure \ref{fig:helmholtz} we also show the norms of each block of calculated coefficients for the approximation of the solution to the inhomogeneous variable-coefficient Helmholtz equation with various $k$ values. Here, we use $N = 100$, that is, $(N+1)^2 = 10,201$ unknowns. Once again, the rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution, and we see that we achieve spectral convergence.

@@

Find $u(x,y,z)$ given a function $f(x,y,z)$ such that:
\begin{align}
	\begin{cases}
    		\DeltaS^2 u(x,y,z) = f(x,y,z) &\quad \text{in } \Omega \\
		u(x,y,z) = 0, \quad \frac{\partial u}{\partial n}(x,y,z) = \grad_S \: u(x,y,z) \cdot \unitvec{n}(x,y,z) = 0 &\quad \text{on } \partial \Omega
	\end{cases}
	\label{eqn:biharmonic}
\end{align}
where $\DeltaS^2$ is the Biharmonic operator, noting the imposition of zero Dirichlet and Neumann boundary conditions on $u$. For clarity, we reiterate that the unit normal vector in this sense is simply $\unitvec{n}(x,y,z) = \unitvec{n}(\vec x) := {\vec x \over \norm{\vec x}} = \vec x$ (see Section \ref{subsection:furtherdiffoperators}). In Figure \ref{fig:biharmonic} we see the solution to the Biharmonic equation (\ref{eqn:biharmonic}) in the spherical cap $\Omega$. In Figure \ref{fig:biharmonic} we also show the norms of each block of calculated coefficients of the approximation for four more complex right-hand sides of the biharmonic equation with $N = 100$, that is, $(N+1)^2 = 10,201$ unknowns. Once again, the rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution, and we see that we achieve algebraic convergence for these more complex functions.

>>>>

\bstodoinline{You mentioned before that you didnt like the use of $i$ being an index throughout (as in $n,k,i$). Is this still the case?}

@@

Define $X := (J_x^{(0)})^\top$, $Y := (J_y^{(0)})^\top$, $Z := (J_z^{(0)})^\top$. We can obtain the matrix operator for the variable-coefficient function $v(x,y,z)$ by using the Clenshaw algorithm with matrix inputs as the Jacobi matrices $X, Y, Z$, yielding an operator matrix of the same dimension as the input Jacobi matrices a la the procedure introduced in \cite{olver2019triangle}. We can denote the resulting operator acting on coefficients in the $\bigscopNo$ space by $v(X, Y, Z)$. In matrix-vector notation, our system hence becomes:
\begin{align*}
    (\Delta_W^{(1)} + k^2 \:T^{(0)\to(1)} \: V \: T_W^{(1)\to(0)}) \: \mathbf{u} = \mathbf{f}
\end{align*}
which can be solved to find $\mathbf{u}$. We can see the sparsity and structure of this matrix system in Figure \ref{fig:sparsity} with $v(x,y,z) = zxy^2$ as an example. In Figure \ref{fig:helmholtz} we see the solution to the inhomogeneous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the spherical cap $\Omega$, with $f(x,y,z) = y e^x \genjacw^{(1,0)}(z)$, $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$ and $k = 100$. In Figure \ref{fig:helmholtz} we also show the norms of each block of calculated coefficients for the approximation of the solution to the inhomogeneous variable-coefficient Helmholtz equation with various $k$ values. Here, we use $N = 200$, that is, $(N+1)^2 = 40,401$ unknowns. Once again, the rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution, and we see that we achieve spectral convergence.

@@

Find $u(x,y,z)$ given a function $f(x,y,z)$ such that:
\begin{align}
	\begin{cases}
    		\DeltaS^2 u(x,y,z) = f(x,y,z) &\quad \text{in } \Omega \\
		u(x,y,z) = 0, \quad \frac{\partial u}{\partial n}(x,y,z) = \grad_S \: u(x,y,z) \cdot \unitvec{n}(x,y,z) = 0 &\quad \text{on } \partial \Omega
	\end{cases}
	\label{eqn:biharmonic}
\end{align}
where $\DeltaS^2$ is the Biharmonic operator, noting the imposition of zero Dirichlet and Neumann boundary conditions on $u$. For clarity, we reiterate that the unit normal vector in this sense is simply $\unitvec{n}(x,y,z) = \unitvec{n}(\vec x) := {\vec x \over \norm{\vec x}} = \vec x$ (see Section \ref{subsection:furtherdiffoperators}). In Figure \ref{fig:biharmonic} we see the solution to the Biharmonic equation (\ref{eqn:biharmonic}) in the spherical cap $\Omega$. In Figure \ref{fig:biharmonic} we also show the norms of each block of calculated coefficients of the approximation for four more complex right-hand sides of the biharmonic equation with $N = 200$, that is, $(N+1)^2 = 40,401$ unknowns. Once again, the rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution, and we see that we achieve algebraic convergence for these more complex functions.
